Perturbation Theory in the Complex Plane: Exceptional Points and Where to Find Them

### 2020-12-02 Manuscript/EPAWTFT.tex
Quadratic section

%==========================================%
\subsection{Quadratic Approximant}
%==========================================%
Quadratic approximants \hugh{are designed} to model the singularity structure of the energy 
function $E(\lambda)$ via a generalised version of the square-root singularity 
expression \cite{Mayer_1985,Goodson_2011,Goodson_2019}
\begin{equation}
	\label{eq:QuadApp}
	E(\lambda) = \frac{1}{2 Q(\lambda)} \qty[ P(\lambda) \pm \sqrt{P^2(\lambda) - 4 Q(\lambda) R(\lambda)} ]
\end{equation}
with the polynomials 
\begin{align}
	\label{eq:PQR}
	P(\lambda) & = \sum_{k=0}^{d_P} p_k \lambda^k,
	&
	Q(\lambda) & = \sum_{k=0}^{d_Q} q_k \lambda^k, 
	&
	R(\lambda) & = \sum_{k=0}^{d_R} r_k \lambda^k 
\end{align}
defined such that $d_P + d_Q + d_R = n - 1$, and $n$ is the truncation order of the Taylor series of $E(\lambda)$.
Recasting Eq.~\eqref{eq:QuadApp} as a second-order expression in $E(\lambda)$, \ie,
\begin{equation}
	Q(\lambda) E^2(\lambda) - P(\lambda) E(\lambda) + R(\lambda) \sim \order*{\lambda^{n+1}}
\end{equation}
and substituting $E(\lambda$) by its $n$th-order expansion and the polynomials by 
their respective expressions \eqref{eq:PQR} yields $n+1$ linear equations for the coefficients 
$p_k$, $q_k$, and $r_k$ (where we are free to assume that $q_0 = 1$).
A quadratic approximant, characterised by the label $[d_P/d_Q,d_R]$, generates, by construction, 
$n_\text{bp} = \max(2d_p,d_q+d_r)$ branch points at the roots of the polynomial $P^2(\lambda) - 4 Q(\lambda) R(\lambda)$ \hugh{and $d_q$ poles at the roots of $Q(\lambda)$}.

Generally, the diagonal sequence of quadratic approximant, 
\ie, $[0/0,0]$, $[1/0,0]$, $[1/0,1]$, $[1/1,1]$, $[2/1,1]$, 
is of particular interest \hugh{as the order of the corresponding Taylor series increases on each step.
However, while a quadratic approximant can reproduce multiple branch points, it can only describe 
a total of two branches.
Since every branch point must therefore correspond to a degeneracy of the same two branches, this constraint
can hamper the faithful description of more complicated singularity structures such as the MP energy surface.
Despite this limitiation,} Ref.~\onlinecite{Goodson_2000a} demonstrates that quadratic approximants 
provide convergent results in the most divergent cases considered by Olsen and 
collaborators\cite{Christiansen_1996,Olsen_1996} 
and Leininger \etal \cite{Leininger_2000}

@@

\begin{table}[b]
    \caption{Estimate \hugh{for the distance of the closest singularity (pole or branch point) to the origin $\abs{\lc}$}
    in the UMP energy function provided 
    by various resummation techniques at $U/t = 3$ and $7$.
	The truncation degree of the Taylor expansion $n$ of $E(\lambda)$ and the number of branch 
    points $n_\text{bp} = \max(2d_p,d_q+d_r)$ generated by the quadratic approximants are also reported.
	\label{tab:QuadUMP}}
	\begin{ruledtabular}
		\begin{tabular}{lccccccc}
            &			&			&					&	\mc{2}{c}{$\abs{\lc}$}	&	\mc{2}{c}{$E_{-}(\lambda = 1)$}			\\
																		\cline{5-6}\cline{7-8}
			\mc{2}{c}{Method}		&	$n$		&	$n_\text{bp}$	&	$U/t = 3$	&	$U/t = 7$	&	$U/t = 3$	&	$U/t = 7$	\\
			\hline
			Taylor		&	     	&	2		&					&	    		&	    		&	$-0.74074$	&	$-0.29155$	\\
                        &	     	&	3		&					&	    		&	    		&	$-0.78189$	&	$-0.29690$	\\
			         	&	     	&	4		&					&	    		&	    		&	$-0.82213$	&	$-0.30225$	\\
						&	     	&	5		&					&	    		&	    		&	$-0.85769$	&	$-0.30758$	\\
                        &	     	&   6		&					&	    		&	    		&	$-0.88882$	&	$-0.31289$	\\
			\hline
			Pad\'e		&	[1/1]	&	2		&					&	$9.000$		&	$49.00$		&	$-0.75000$	&	$-0.29167$	\\
                        &	[2/2]	&	4		&					&	$0.974$		&	$1.003$		&	$\hphantom{-}0.75000$	&	$-17.9375$	\\
			         	&	[3/3]	&	6		&					&	$1.141$		&	$1.004$		&	$-1.10896$	&	$-1.49856$	\\
						&	[4/4]	&	8		&					&	$1.068$		&	$1.003$		&	$-0.85396$	&	$-0.33596$	\\
                        &	[5/5]	&	10		&					&	$1.122$		&	$1.004$		&	$-0.97254$	&	$-0.35513$	\\
            \hline
			Quadratic	&	[2/1,2]	&	6		&	4				&	$1.086$		&	$1.003$		&	$-1.01009$	&	$-0.53472$	\\
						&	[2/2,2]	&	7		&	4				&	$1.082$		&	$1.003$		&	$-1.00553$	&	$-0.53463$	\\
						&	[3/2,2]	&	8		&	6				&	$1.082$		&	$1.001$		&	$-1.00568$	&	$-0.52473$	\\
						&	[3/2,3]	&	9		&	6				&	$1.071$		&	$1.002$		&	$-0.99973$	&	$-0.53102$	\\
                        &	[3/3,3]	&	10		&	6				&	$1.071$		&	$1.002$		&	$-0.99966$	&	$-0.53103$	\\[0.5ex]
            (pole-free)	&	[3/0,2]	&	6		&	6				&	$1.059$	    &	$1.003$ 	&	$-1.13712$	&	$-0.57199$	\\
		                &	[3/0,3]	&	7		&	6				&	$1.073$		&	$1.002$		&	$-1.00335$	&	$-0.53113$	\\
						&	[3/0,4]	&	8		&	6				&	$1.071$		&	$1.002$		&	$-1.00074$	&	$-0.53116$	\\
						&	[3/0,5]	&	9		&	6				&	$1.070$		&	$1.002$		&	$-1.00042$	&	$-0.53114$	\\
						&	[3/0,6]	&	10		&	6				&	$1.070$		&	$1.002$		&	$-1.00039$	&	$-0.53113$	\\
			\hline
			Exact		&			&			&					&	$1.069$		&	$1.002$		&	$-1.00000$	&	$-0.53113$	\\
		\end{tabular}
	\end{ruledtabular}
\end{table}

\begin{figure*}
    \begin{subfigure}{0.32\textwidth}
	\includegraphics[height=0.85\textwidth]{ump_qa322}	
    \subcaption{\label{subfig:ump_ep_to_cp} [3/2,2] Quadratic}
    \end{subfigure}
    %
    \begin{subfigure}{0.32\textwidth}
	\includegraphics[height=0.85\textwidth]{ump_exact}
		\subcaption{\label{subfig:ump_cp_surf} Exact}
    \end{subfigure}
    %
	\begin{subfigure}{0.32\textwidth}
        \includegraphics[height=0.85\textwidth]{ump_qa304}	
    \subcaption{\label{subfig:ump_cp} [3/0,4] Quadratic}
    \end{subfigure}
\caption{%
\hugh{% 
Comparison of the [3/2,2] and [3/0,4] quadratic approximants with the exact UMP energy surface in the complex $\lambda$ 
plane with $U/t = 3$. 
Both quadratic approximants correspond to the same truncation degree of the Taylor series and model the branch points 
using a radicand polynomial of the same order.
However, the [3/2,2] approximant introduces poles into the surface that limits it accuracy, while the [3/0,4] approximant
is free of poles.}
\label{fig:nopole_quad}}
\end{figure*}

For the RMP series of the Hubbard dimer, the $[0/0,0]$ and $[1/0,0]$ quadratic approximant 
are quite poor approximations, but the $[1/0,1]$ version perfectly models the RMP energy
function by predicting a single pair of EPs at $\lambda_\text{EP} = \pm \i 4t/U$.
This is expected from the form of the RMP energy [see Eq.~\eqref{eq:E0MP}], which matches 
the ideal target for quadratic approximants.
\hugh{%
Furthermore, the greater flexibility of the diagonal quadratic approximants provides a significantly
improved model of the UMP energy in comparison to the Pad\'e approximants or Taylor series.
In particular, these quadratic approximants provide an effective model for the avoided crossings 
(Fig.~\ref{fig:QuadUMP}) and an improved estimate for the distance of the 
closest branch point  to the origin.
Table~\ref{tab:QuadUMP} shows that they provide remarkably accurate 
estimates of the ground-state energy at $\lambda = 1$.}

\hugh{%
While the diagonal quadratic approximants provide significanty improved estimates of the 
ground-state energy, we can use our knowledge of the UMP singularity structure to develop
even more accurate results.
We have seen in previous sections that the UMP energy surface
contains only square-root branch cuts that approach the real axis in the limit $U/t \rightarrow \infty$. 
Since there are no true poles on this surface, we can obtain more accurate quadratic approximants by
taking $d_q = 0$ and increasing $d_r$ to retain equivalent accuracy in the square-root term.
Fig.~\ref{fig:nopole_quad} illustrates this improvement for the pole-free [3/0,4] quadratic
approximant compared to the [3/2,2] approximant with the same truncation degree in the Taylor
expansion.
Clearly modelling the square-root branch point using $d_q = 2$ has the negative effect of
introducing spurious poles in the energy, while focussing purely on the branch point with $d_q = 0$
leads to a significantly improved model.
Table~\ref{tab:QuadUMP} shows that these pole-free quadratic approximants
provide a rapidly convergent series with essentially exact energies at low order.
}

\hugh{Finally, to emphasise the improvement that can be gained by using either Pad\'e, diagonal quadratic,
or pole-free approximants, we consider the energy and error obtained using only the first 10 terms in the UMP
in Table~\ref{tab:UMP_order10}.
The accuracy of these approximants reinforces how our understanding of the MP
energy surface in the complex plane can be leveraged to significantly improve estimates of the exact
energy using low-order perturbation expansions.
}

\begin{table}[h]
	\caption{
    \hugh{%
    Estimate and associated error of the exact UMP energy at $U/t = 7$ for 
    various approximants using up to ten terms in the Taylor expansion.
    }
	\label{tab:UMP_order10}}
	\begin{ruledtabular}
		\begin{tabular}{lccc}
            \mc{2}{c}{Method}	 &	$E_{-}(\lambda = 1)$ & Abs.\ Error \\ 
			\hline
            Taylor		         &	 10	         &	$-0.33338$      &  $0.197290$ \\
            Pad\'e               &   [5/5]       &  $-0.35513$      &  $0.176000$ \\
            Quadratic (diagonal) &   [3/3,3]     &  $-0.53104$      &  $0.000103$ \\
            Quadratic (pole-free)&   [3/0,6]     &  $-0.53113$      &  $0.000003$ \\
			\hline
            Exact		         &	        	 &	$-0.53113$	     &            \\
		\end{tabular}
	\end{ruledtabular}
\end{table}

@@

While the Pad\'e and quadratic approximants can yield a convergent series representation
in cases where the standard MP series diverges, there is no guarantee that the rate of convergence
will be fast enough for low-order approximations to be useful.
However, these low-order partial sums or approximants often contain a remarkable amount of information
that can be used to extract further information about the exact result.
The Shanks transformation presents one approach for extracting this information
and accelerating the rate of convergence of a sequence.\cite{Shanks_1955}

>>>>

%==========================================%
\subsection{Quadratic Approximant}
%==========================================%
Quadratic approximants are designed to model the singularity structure of the energy 
function $E(\lambda)$ via a generalised version of the square-root singularity 
expression \cite{Mayer_1985,Goodson_2011,Goodson_2019}
\begin{equation}
	\label{eq:QuadApp}
	E(\lambda) = \frac{1}{2 Q(\lambda)} \qty[ P(\lambda) \pm \sqrt{P^2(\lambda) - 4 Q(\lambda) R(\lambda)} ],
\end{equation}
with the polynomials 
\begin{align}
	\label{eq:PQR}
	P(\lambda) & = \sum_{k=0}^{d_P} p_k \lambda^k,
	&
	Q(\lambda) & = \sum_{k=0}^{d_Q} q_k \lambda^k, 
	&
	R(\lambda) & = \sum_{k=0}^{d_R} r_k \lambda^k,
\end{align}
defined such that $d_P + d_Q + d_R = n - 1$, and $n$ is the truncation order of the Taylor series of $E(\lambda)$.
Recasting Eq.~\eqref{eq:QuadApp} as a second-order expression in $E(\lambda)$, \ie,
\begin{equation}
	Q(\lambda) E^2(\lambda) - P(\lambda) E(\lambda) + R(\lambda) \sim \order*{\lambda^{n+1}},
\end{equation}
and substituting $E(\lambda$) by its $n$th-order expansion and the polynomials by 
their respective expressions \eqref{eq:PQR} yields $n+1$ linear equations for the coefficients 
$p_k$, $q_k$, and $r_k$ (where we are free to assume that $q_0 = 1$).
A quadratic approximant, characterised by the label $[d_P/d_Q,d_R]$, generates, by construction, 
$n_\text{bp} = \max(2d_p,d_q+d_r)$ branch points at the roots of the polynomial 
$P^2(\lambda) - 4 Q(\lambda) R(\lambda)$ and $d_q$ poles at the roots of $Q(\lambda)$.

Generally, the diagonal sequence of quadratic approximant, 
\ie, $[0/0,0]$, $[1/0,0]$, $[1/0,1]$, $[1/1,1]$, $[2/1,1]$, 
is of particular interest as the order of the corresponding Taylor series increases on each step.
However, while a quadratic approximant can reproduce multiple branch points, it can only describe 
a total of two branches.
\titou{Since every branch points must therefore correspond to a degeneracy of the same two branches,} this constraint
can hamper the faithful description of more complicated singularity structures such as the MP energy surface.
Despite this limitation, Ref.~\onlinecite{Goodson_2000a} demonstrates that quadratic approximants 
provide convergent results in the most divergent cases considered by Olsen and 
collaborators\cite{Christiansen_1996,Olsen_1996} 
and Leininger \etal \cite{Leininger_2000}

@@

\begin{table}[b]
    \caption{Estimate for the distance of the closest singularity (pole or branch point) to the origin $\abs{\lc}$
    in the UMP energy function provided by various resummation techniques at $U/t = 3$ and $7$.
	The truncation degree of the Taylor expansion $n$ of $E(\lambda)$ and the number of branch 
    points $n_\text{bp} = \max(2d_p,d_q+d_r)$ generated by the quadratic approximants are also reported.
	\label{tab:QuadUMP}}
	\begin{ruledtabular}
		\begin{tabular}{lccccccc}
            &			&			&					&	\mc{2}{c}{$\abs{\lc}$}	&	\mc{2}{c}{$E_{-}(\lambda = 1)$}			\\
																		\cline{5-6}\cline{7-8}
			\mc{2}{c}{Method}		&	$n$		&	$n_\text{bp}$	&	$U/t = 3$	&	$U/t = 7$	&	$U/t = 3$	&	$U/t = 7$	\\
			\hline
			Taylor		&	     	&	2		&					&	    		&	    		&	$-0.74074$	&	$-0.29155$	\\
                        &	     	&	3		&					&	    		&	    		&	$-0.78189$	&	$-0.29690$	\\
			         	&	     	&	4		&					&	    		&	    		&	$-0.82213$	&	$-0.30225$	\\
						&	     	&	5		&					&	    		&	    		&	$-0.85769$	&	$-0.30758$	\\
                        &	     	&   6		&					&	    		&	    		&	$-0.88882$	&	$-0.31289$	\\
			\hline
			Pad\'e		&	[1/1]	&	2		&					&	$9.000$		&	$49.00$		&	$-0.75000$	&	$-0.29167$	\\
                        &	[2/2]	&	4		&					&	$0.974$		&	$1.003$		&	$\hphantom{-}0.75000$	&	$-17.9375$	\\
			         	&	[3/3]	&	6		&					&	$1.141$		&	$1.004$		&	$-1.10896$	&	$-1.49856$	\\
						&	[4/4]	&	8		&					&	$1.068$		&	$1.003$		&	$-0.85396$	&	$-0.33596$	\\
                        &	[5/5]	&	10		&					&	$1.122$		&	$1.004$		&	$-0.97254$	&	$-0.35513$	\\
            \hline
			Quadratic	&	[2/1,2]	&	6		&	4				&	$1.086$		&	$1.003$		&	$-1.01009$	&	$-0.53472$	\\
						&	[2/2,2]	&	7		&	4				&	$1.082$		&	$1.003$		&	$-1.00553$	&	$-0.53463$	\\
						&	[3/2,2]	&	8		&	6				&	$1.082$		&	$1.001$		&	$-1.00568$	&	$-0.52473$	\\
						&	[3/2,3]	&	9		&	6				&	$1.071$		&	$1.002$		&	$-0.99973$	&	$-0.53102$	\\
                        &	[3/3,3]	&	10		&	6				&	$1.071$		&	$1.002$		&	$-0.99966$	&	$-0.53103$	\\[0.5ex]
            (pole-free)	&	[3/0,2]	&	6		&	6				&	$1.059$	    &	$1.003$ 	&	$-1.13712$	&	$-0.57199$	\\
		                &	[3/0,3]	&	7		&	6				&	$1.073$		&	$1.002$		&	$-1.00335$	&	$-0.53113$	\\
						&	[3/0,4]	&	8		&	6				&	$1.071$		&	$1.002$		&	$-1.00074$	&	$-0.53116$	\\
						&	[3/0,5]	&	9		&	6				&	$1.070$		&	$1.002$		&	$-1.00042$	&	$-0.53114$	\\
						&	[3/0,6]	&	10		&	6				&	$1.070$		&	$1.002$		&	$-1.00039$	&	$-0.53113$	\\
			\hline
			Exact		&			&			&					&	$1.069$		&	$1.002$		&	$-1.00000$	&	$-0.53113$	\\
		\end{tabular}
	\end{ruledtabular}
\end{table}

\begin{figure*}
    \begin{subfigure}{0.32\textwidth}
	\includegraphics[height=0.85\textwidth]{ump_qa322}	
    \subcaption{\label{subfig:ump_ep_to_cp} [3/2,2] Quadratic}
    \end{subfigure}
    %
    \begin{subfigure}{0.32\textwidth}
	\includegraphics[height=0.85\textwidth]{ump_exact}
		\subcaption{\label{subfig:ump_cp_surf} Exact}
    \end{subfigure}
    %
	\begin{subfigure}{0.32\textwidth}
        \includegraphics[height=0.85\textwidth]{ump_qa304}	
    \subcaption{\label{subfig:ump_cp} [3/0,4] Quadratic}
    \end{subfigure}
\caption{%
Comparison of the [3/2,2] and [3/0,4] quadratic approximants with the exact UMP energy surface in the complex $\lambda$ 
plane with $U/t = 3$. 
Both quadratic approximants correspond to the same truncation degree of the Taylor series and model the branch points 
using a radicand polynomial of the same order.
However, the [3/2,2] approximant introduces poles into the surface that limits it accuracy, while the [3/0,4] approximant
is free of poles.}
\label{fig:nopole_quad}
\end{figure*}

For the RMP series of the Hubbard dimer, the $[0/0,0]$ and $[1/0,0]$ quadratic approximant 
are quite poor approximations, but the $[1/0,1]$ version perfectly models the RMP energy
function by predicting a single pair of EPs at $\lambda_\text{EP} = \pm \i 4t/U$.
This is expected from the form of the RMP energy [see Eq.~\eqref{eq:E0MP}], which matches 
the ideal target for quadratic approximants.
Furthermore, the greater flexibility of the diagonal quadratic approximants provides a significantly
improved model of the UMP energy in comparison to the Pad\'e approximants or Taylor series.
In particular, these quadratic approximants provide an effective model for the avoided crossings 
(Fig.~\ref{fig:QuadUMP}) and an improved estimate for the distance of the 
closest branch point  to the origin.
Table~\ref{tab:QuadUMP} shows that they provide remarkably accurate 
estimates of the ground-state energy at $\lambda = 1$.

While the diagonal quadratic approximants provide significanty improved estimates of the 
ground-state energy, we can use our knowledge of the UMP singularity structure to develop
even more accurate results.
We have seen in previous sections that the UMP energy surface
contains only \titou{square-root branch cuts} that approach the real axis in the limit $U/t \to \infty$. 
Since there are no true poles on this surface, we can obtain more accurate quadratic approximants by
taking $d_q = 0$ and increasing $d_r$ to retain equivalent accuracy in the square-root term [see Eq.\eqref{eq:QuadApp}].
Figure~\ref{fig:nopole_quad} illustrates this improvement for the pole-free [3/0,4] quadratic
approximant compared to the [3/2,2] approximant with the same truncation degree in the Taylor
expansion.
Clearly modelling the square-root branch point using $d_q = 2$ has the negative effect of
introducing spurious poles in the energy, while focussing purely on the branch point with $d_q = 0$
leads to a significantly improved model.
Table~\ref{tab:QuadUMP} shows that these pole-free quadratic approximants
provide a rapidly convergent series with essentially exact energies at low order.

Finally, to emphasise the improvement that can be gained by using either Pad\'e, diagonal quadratic,
or pole-free approximants, we consider the energy and error obtained using only the first 10 terms of the UMP series
in Table~\ref{tab:UMP_order10}.
The accuracy of these approximants reinforces how our understanding of the MP
energy surface in the complex plane can be leveraged to significantly improve estimates of the exact
energy using low-order perturbation expansions.

\begin{table}[h]
	\caption{
    Estimate and associated error of the exact UMP energy at $U/t = 7$ for 
    various approximants using up to ten terms in the Taylor expansion.
	\label{tab:UMP_order10}}
	\begin{ruledtabular}
		\begin{tabular}{lccc}
            \mc{2}{c}{Method}	 &	$E_{-}(\lambda = 1)$ & Abs.\ Error \\ 
			\hline
            Taylor		         &	 10	         &	$-0.33338$      &  $0.197290$ \\
            Pad\'e               &   [5/5]       &  $-0.35513$      &  $0.176000$ \\
            Quadratic (diagonal) &   [3/3,3]     &  $-0.53104$      &  $0.000103$ \\
            Quadratic (pole-free)&   [3/0,6]     &  $-0.53113$      &  $0.000003$ \\
			\hline
            Exact		         &	        	 &	$-0.53113$	     &            \\
		\end{tabular}
	\end{ruledtabular}
\end{table}

@@

While the Pad\'e and quadratic approximants can yield a convergent series representation
in cases where the standard MP series diverges, there is no guarantee that the rate of convergence
will be fast enough for low-order approximations to be useful.
However, these low-order partial sums or approximants often contain a remarkable amount of information
that can be used to extract further information about the exact result.
The Shanks transformation presents one approach for extracting this information
and accelerating the rate of convergence of a sequence.\cite{Shanks_1955,BenderBook}
